\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{../../lib/approx-scribe}

\usepackage[T1]{fontenc}

\allowdisplaybreaks

\begin{document}

\lecture{11}{Ball-Growing and Multicut}{Antares Chen}{5/10/2019}


% --------------------------------------------------------------------
% INTRODUCTION
% --------------------------------------------------------------------
\input{introduction.tex}


% --------------------------------------------------------------------
% MULTICUT PROBLEM + RELAXATION
% --------------------------------------------------------------------
\input{relaxation.tex}


% --------------------------------------------------------------------
% ROUNDING ALGORITHM + CORRECTNESS
% --------------------------------------------------------------------
\input{rounding.tex}


% --------------------------------------------------------------------
% BALL GROWING
% --------------------------------------------------------------------
\input{ball-growing.tex}


% --------------------------------------------------------------------
% CONCLUSION
% --------------------------------------------------------------------
\input{conclusion.tex}

\bibliographystyle{plain}
\begin{thebibliography}{99}
\bibitem{ARV09}
Arora, S., Rao, S., \& Vazirani, U. (2009). ``Expander flows, geometric embeddings and graph partitioning.'' In \emph{Journal of the ACM (JACM)}, 56(2), 5.

\bibitem{CKKRS06}
Chawla, S., Krauthgamer, R., Kumar, R., Rabani, Y., \& Sivakumar, D. (2006). ``On the hardness of approximating multicut and sparsest-cut.'' In \emph{computational complexity}, 15(2), 94-114.

\bibitem{GVY96}
Garg, N., Vazirani, V. V., \& Yannakakis, M. (1996). ``Approximate max-flow min-(multi) cut theorems and their applications.'' In \emph{SIAM Journal on Computing}, 25(2), 235-251.

\bibitem{KM09}
Kelner, J. A., \& Madry, A. (2009, October). ``Faster generation of random spanning trees''. In \emph{2009 50th Annual IEEE Symposium on Foundations of Computer Science} (pp. 13-21). IEEE.

\bibitem{LR88}
Leighton, T., \& Rao, S. (1988). ``An approximate max-flow min-cut theorem for uniform multicommodity flow problems with applications to approximation algorithms.'' In \emph{[Proceedings 1988] 29th Annual Symposium on Foundations of Computer Science}. (pp. 422-431). IEEE.

\bibitem{LR99}
Leighton, T., \& Rao, S. (1999). ``Multicommodity max-flow min-cut theorems and their use in designing approximation algorithms.'' In \emph{Journal of the ACM (JACM)}, 46(6), 787-832.

\bibitem{Tre05}
Trevisan, L. (2005). ``Approximation algorithms for unique games.'' In \emph{46th Annual IEEE Symposium on Foundations of Computer Science} (pp. 197-205). IEEE.
\end{thebibliography}

\end{document}
